The Success of College Basketball Teams can be predicted using several advanced statistics

### 2022-04-27 Paper-On-NCAAB-Stats/inputs/data/datasheet_template.Rmd
Final Paper

---
title: "Datasheet for dataset used in 'The Success of College Basketball Teams can be predicted using several advanced team statistics'"
author: 
  - Thomas D'Onofrio
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references_datasheet.bib
---

>>>>

---
title: "Datasheet for dataset used in 'The Success of College Basketball Teams can be predicted using several advanced team statistics'"
author: 
  - Thomas D'Onofrio
thanks: "Code and data are available at: https://github.com/TDonofrio62/Predicting-College-Basketball-Success."
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references_datasheet.bib
---

### 2022-04-27 Paper-On-NCAAB-Stats/outputs/paper/Paper-On-NCAAB-Stats.Rmd
Final Paper

---
title: "The Success of College Basketball Teams can be predicted using several advanced team statistics"
subtitle: "A Linear Model Approach"
author: 
  - Thomas D'Onofrio
thanks: "Code and data are available at: https://github.com/TDonofrio62/Paper-On-NCAAB-Stats."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The March Madness college basketball tournament is one of the most exciting sports events in the world due to the intense games, unpredictable results, and the creation of brackets by millions. In this paper, data displaying each college basketball schools’ advanced team statistics was used to create a linear model with Simple Rating System as the response variable, where the explanatory variables are the team statistics that play the largest factor in leading to team success. It was found that true shooting percentage, total rebound percentage, and turnover percentage are the most important team statistics that affect the amount of success a team has, meaning fans can use these statistics to fill out their brackets more successfully. Unfortunately, the data is slightly biased due to inconsistent scorekeeping across teams and arenas, and the racial biases of referees when calling fouls, causing some implications to the findings."
output:
  bookdown::pdf_document2:
    fig_caption: true
toc: FALSE
bibliography: references.bib
---

@@

We can plug all these coefficients into the $\beta's$ in the formula explained in Section \@ref(model) to get the completed formula for our final model, $\hat{y_i} = -98.62 + 0.92x_1 + 1.63x_2 - 2.04x_3$. Doing so makes our model much easier to interpret. In order to estimate what a team's simple rating system value would be, you plug in their true shooting percentage for $x_1$, their total rebound percentage for $x_2$, and their turnover percentage for $x_3$, and the output of $\hat{y_i}$ would be their rating. For example, if a team has a true shooting percentage of 50%, a total rebounding percentage of 40%, and a turnover percentage of 20%, the model estimates that their simple rating system would be $\hat{y_i} = -98.62 + 0.92(50) + 1.63(40) - 2.04(20) = -28.22$. This shows how the model can successfully predict a team's success with the use of advanced team statistics such as our predictor variables that were used, answering the research question.

@@

```{r 3d, echo=FALSE, fig.cap= "4D plot showing the relationship between the models response variable, Simple Rating System, and its three predictor variables"}

@@

scatter3D(ncaab_stats_train$ts_percent, ncaab_stats_train$trb_percent, ncaab_stats_train$srs, 
          colvar = ncaab_stats_train$tov_percent, 
          bty = 'b2', pch = 19, cex = 0.5, 
          xlab = "TS%", ylab = "TRB%", zlab = "SRS", 
          clab = "TOV%", ticktype = "detailed")
```

@@

```{r assumplots, echo=FALSE, fig.height=8, fig.cap="Residuals and Fitted Value plots for linear regression model predicting Simple Rating System using a training dataset on college basketball statistics"}

@@

# Appendix {-}

>>>>

---
title: "The Success of College Basketball Teams can be predicted using several advanced statistics"
subtitle: "A Linear Regression Approach"
author: 
  - Thomas D'Onofrio
thanks: "Code and data are available at: https://github.com/TDonofrio62/Predicting-College-Basketball-Success."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "The March Madness college basketball tournament is one of the most exciting sports events in the world due to the intense games, unpredictable results, and the creation of brackets by millions. In this paper, data displaying each college basketball schools’ advanced team statistics was used to create a linear model with Simple Rating System as the response variable, where the explanatory variables are the team statistics that play the largest factor in leading to team success. It was found that true shooting percentage, total rebound percentage, and turnover percentage are the most important team statistics that affect the amount of success a team has, meaning fans can use these statistics to fill out their brackets more successfully. Unfortunately, the data is slightly biased due to inconsistent scorekeeping across teams and arenas, and the racial biases of referees when calling fouls, causing some implications to the findings."
output:
  bookdown::pdf_document2:
    fig_caption: true
toc: FALSE
bibliography: references.bib
---

@@

We can plug all these coefficients into the $\beta's$ in the formula explained in Section \@ref(model) to get the completed formula for our final model, $$\hat{y_i} = -98.62 + 0.92x_1 + 1.63x_2 - 2.04x_3$$. Doing so makes our model much easier to interpret. In order to estimate what a team's simple rating system value would be, you plug in their true shooting percentage for $x_1$, their total rebound percentage for $x_2$, and their turnover percentage for $x_3$, and the output of $\hat{y_i}$ would be their rating. For example, if a team has a true shooting percentage of 50%, a total rebounding percentage of 40%, and a turnover percentage of 20%, the model estimates that their simple rating system would be $\hat{y_i} = -98.62 + 0.92(50) + 1.63(40) - 2.04(20) = -28.22$. This shows how the model can successfully predict a team's success with the use of advanced team statistics such as our predictor variables that were used, answering the research question.

@@

```{r 3d, echo=FALSE, fig.height=5, fig.width=6, fig.cap= "4D plot showing the relationship between the models response variable, Simple Rating System, and its three predictor variables"}

@@

scatter3D(ncaab_stats_train$ts_percent, ncaab_stats_train$trb_percent, ncaab_stats_train$srs, 
          colvar = ncaab_stats_train$tov_percent, 
          bty = 'b2', pch = 19, cex = 0.5, 
          xlab = "TS%", ylab = "TRB%", zlab = "SRS", 
          clab = "TOV%", ticktype = "detailed")

@@

```{r assumplots, echo=FALSE, fig.height=7, fig.cap="Residuals and Fitted Value plots for linear regression model predicting Simple Rating System using a training dataset on college basketball statistics"}

@@

# Appendix
